### 2020-07-26 inst/sample_size_estimator/www/example.Rmd
added shiny app to package with rank and protein selection

<html>
<div class = "row">
<div class = "col-md-6">
<h3>Protein Abundance File</h3>
<p>This must be a csv file in the <b>wide format</b>. Rows are proteins, columns are 
    Biological replicates or samples, cells are protein  abundances. The following
    image is an example, example data file can be found <a href = "https://github.com/Vitek-Lab/MSstatsSampleSize-Shiny/blob/master/OV_SRM_train.csv " target="_blank"><b>here</b></a>
    </p>
<img src = ab.PNG width = "100%" height = "350">

</div>

<div class = "col-md-6">
<h3>Sample Annotation File</h3>
<p> This must be a csv file in the <b>long format</b>. It includes columns BioReplicate
    for sample ID and Condition for group information. BioReplicate information
    should be the same as the column names of Preliminary data. The following
    image is an example, example data file can be found <a href = "https://github.com/Vitek-Lab/MSstatsSampleSize-Shiny/blob/master/OV_SRM_train_annotation.csv" target="_blank"><b>here</b></a>
    </p>
<img src = an.PNG  height = "350">
</div>
</div>
</html>